---
title: "Manuscript draft"
author: "Molly Wilson"
date: "Dec. 2018"
output:
  html_document:
    toc: true
    toc_float: true
    theme: united
    number_sections: true
---

```{r, echo=F, message=F, warning=F}
library(startR)
library(here)
library(tidyverse)
library(gridExtra)
library(ggfortify) # autoplot(prcomp())
library(mgcv)
library(sme)

sum_site <- read.csv(here("data","sum_site.csv"), stringsAsFactors=F) %>%
  select(-X)
sum_id <- read.csv(here("data","sum_id.csv"), stringsAsFactors=F) %>%
  select(-X) %>% na.omit()
sum_ss <- read.csv(here("data","sum_site_sp.csv"), stringsAsFactors=F) %>%
  select(-X)
sum_ssp <- read.csv(here("data","sum_site_sp_ph.csv"), stringsAsFactors=F) %>%
  select(-X)
```

# Intro + motivation

* Grazing by herbivorous fish is an established driver of coral reef health/dynamics.
* As reefs become increasingly degraded/threatened, there is growing interest in understanding how much herbivory is needed to sustain coral reef health (e.g. Karr et al. 2015, McClanahan et al. 2011).
* However, not all herbivores are the same in terms of their grazing impact…
    * Recent developments in functional classifications, difference across species in terms of selectivity, etc. (e.g. Adam et al. 2018).
* In addition to these inter-species differences in functional role, differences in herbivore behavior may also affect grazing impact and may drive variation in herbivore function among individuals or across reef systems.
* Previous work has shown that herbivore behavior is sensitive to...
    * Predator presence (and reef structure) (Catano, Madin)
    * Presence of other herbivores (Gil)
    * Resource availability (Carlson, Tootell)
* But there is very limited data on how herbivore feeding behaviors might vary across reef systems, as well as the environmental factors associated with grazing differences.
* Here we document variability in multiple components of scarid feeding behavior across reef sites in three Caribbean islands and explore potential underlying mechanisms, ecological impacts, and management implications.
* We demonstrate a need for further experimental and modelling work to unpack these drivers and understand their ecological consequences.

# Methods

## Data collection
* Sites (13, [map]), timeline
* Fish and benthic community data
* Behavioral data:
    - Species: *Sparisoma viride* and *Scarus vetula*, initial and terminal phase greater than 15cm
    - 2 minute follows, calibration period, outside of previously established FID
    - Recording time budgets as well as bite rates during grazing forays (define)
    - Discarded data of incomplete follows or where fish were visably disturbed/fled
    - Timing of data collection distributed throughout day, 2 dives per site
    - Define behavioral variables
    
## Analysis
* Fish biomass using lenth-weight conversions (appendix)
* PCA of site-level benthic and fish community variables
* GAMM: mgcv, island as random effect, Gaussian distribution, compared using AICc. Used data aggregated to the site level
* (Predicting community grazing impact: calculated length - feeding rate relationships using linear regression for each phase/island/species, applied these to observed site-level fish communities. Integrated bite size data from Bruggemann et al., then compared with predictions made with published species length - feeding rate relationships)

# Results

## Variation in feeding rate across benthic and fish community variables
```{r, echo=F, message=F, warning=F}
q1 <- ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), aes(x = scar_bm/1000, y = fr_mean/60)) +
  geom_smooth(method="loess") +
  geom_point(shape=18) +
  ylim(-10, 60) +
  labs(y = "",
       # y = "S. vetula \nFeeding rate (bites/min)", 
       x = "", 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        plot.margin = margin(0, 1, 0, 1, unit = "pt")
        )

q2 <- ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), aes(x = consp_den, y = fr_mean/60)) +
  geom_smooth(method="loess") +
  geom_point(shape=18) +
  ylim(-10, 60) +
  labs(y = "", 
       x = "",
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(1, 1, 0, 1, unit = "pt")
        )

q3 <- ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), aes(x = rugosity, y = fr_mean/60)) +
  geom_smooth(method="loess") +
  geom_point(shape=18) +
  ylim(-10, 60) +
  labs(y = "", 
       x = "", 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(1, 1, 0, 1, unit = "pt")
        )

q4 <- ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), aes(x = ta_canopy, y = fr_mean/60)) +
  geom_smooth(method="loess") +
  geom_point(shape=18) +
  ylim(-10, 60) +
  labs(y = "", 
       x = "",
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(1, 1, 0, 1, unit = "pt")
        )

q5 <- ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), aes(x = ta_cover, y = fr_mean/60)) +
  geom_smooth(method="loess") +
  geom_point(shape=18) +
  ylim(-10, 60) +
  labs(y = "",
       x = "",
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(1, 1, 0, 1, unit = "pt")
        )

q7 <- ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), aes(x = ma_cover, y = fr_mean/60)) +
  geom_smooth(method="loess") +
  geom_point(shape=18) +
  ylim(-10, 60) +
  labs(y = "", 
       x = "",
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(1, 1, 0, 1, unit = "pt")
        )

s1 <- ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), aes(x = scar_bm/1000, y = fr_mean/60)) +
  geom_smooth(method="loess") +
  geom_point(shape=18) +
  ylim(-5, 20) +
  labs(y = "",
       x = "",
       # y = "S. viride \nFeeding rate (bites/min)", 
       # x = expression(atop(Scarid~biomass, "(kg/100m"^2*")")),
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        plot.margin = margin(0, 1, 0, 1, unit = "pt")
        )

s2 <- ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), aes(x = consp_den, y = fr_mean/60)) +
  geom_smooth(method="loess") +
  geom_point(shape=18) +
  ylim(-5, 20) +
  labs(y = "", 
       x = "",
       # x = expression(atop(Conspec.~density, "(indv./100m"^2*")")),
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(1, 1, 0, 1, unit = "pt")
        )

s3 <- ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), aes(x = rugosity, y = fr_mean/60)) +
  geom_smooth(method="loess") +
  geom_point(shape=18) +
  ylim(-5, 20) +
  labs(y = "", 
       x = "",
       #x = expression("Rugosity"), 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(1, 1, 0, 1, unit = "pt")
        )

s4 <- ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), aes(x = ta_canopy, y = fr_mean/60)) +
  geom_smooth(method="loess") +
  geom_point(shape=18) +
  ylim(-5, 20) +
  labs(y = "", 
       x = "",
       # x = expression("EAM canopy \nheight (mm)"), 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(1, 1, 0, 1, unit = "pt")
        )

s5 <- ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), aes(x = ta_cover, y = fr_mean/60)) +
  geom_smooth(method="loess") +
  geom_point(shape=18) +
  ylim(-5, 20) +
  labs(y = "", 
       x = "",
       # x = expression("EAM percent \ncover"), 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(1, 1, 0, 1, unit = "pt")
        )

s7 <- ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), aes(x = ma_cover, y = fr_mean/60)) +
  geom_smooth(method="loess") +
  geom_point(shape=18) +
  ylim(-5, 20) +
  labs(y = "", 
       x = "",
       # x = expression("Macroalgal \npercent cover"), 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(1, 1, 0, 1, unit = "pt")
        )

grid.arrange(left = "Feeding rate (bites / minute)",
  arrangeGrob(q1, q3, q4, q5, q7, ncol=5, right = "S. vetula"),
  arrangeGrob(
    arrangeGrob(s1, bottom = "Scarid biomass"), 
    arrangeGrob(s3, bottom = "Rugosity"),
    arrangeGrob(s4, bottom = "Turf height"),
    arrangeGrob(s5, bottom = "Turf % cover"),
    arrangeGrob(s7, bottom = "Macro. % cover"),
    ncol=5, right = "S. viride"), 
  nrow=2) 
```

## Variation in feeding intensity across benthic and fish community variables
```{r, echo=F, message=F, warning=F}
q1 <- ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), aes(x = scar_bm/1000, y = for_bites_mean)) +
  geom_smooth(method="loess") +
  geom_point(shape=18) +
  ylim(-10, 25) +
  labs(y = "",
       # y = "S. vetula \nFeeding rate (bites/min)", 
       x = "", 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        plot.margin = margin(1, 1, 0, 1, unit = "pt")
        )

q2 <- ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), aes(x = consp_den, y = for_bites_mean)) +
  geom_smooth(method="loess") +
  geom_point(shape=18) +
  ylim(-10, 25) +
  labs(y = "", 
       x = "",
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(1, 1, 0, 1, unit = "pt")
        )

q3 <- ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), aes(x = rugosity, y = for_bites_mean)) +
  geom_smooth(method="loess") +
  geom_point(shape=18) +
  ylim(-10, 25) +
  labs(y = "", 
       x = "", 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(1, 1, 0, 1, unit = "pt")
        )

q4 <- ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), aes(x = ta_canopy, y = for_bites_mean)) +
  geom_smooth(method="loess") +
  geom_point(shape=18) +
  ylim(-10, 25) +
  labs(y = "", 
       x = "",
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(1, 1, 0, 1, unit = "pt")
        )

q5 <- ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), aes(x = ta_cover, y = for_bites_mean)) +
  geom_smooth(method="loess") +
  geom_point(shape=18) +
  ylim(-10, 25) +
  labs(y = "",
       x = "",
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(1, 1, 0, 1, unit = "pt")
        )

q7 <- ggplot(filter(sum_ssp, species_code == "qup" & phase == "i"), aes(x = ma_cover, y = for_bites_mean)) +
  geom_smooth(method="loess") +
  geom_point(shape=18) +
  ylim(-10, 25) +
  labs(y = "", 
       x = "",
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(1, 1, 0, 1, unit = "pt")
        )

s1 <- ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), aes(x = scar_bm/1000, y = for_bites_mean)) +
  geom_smooth(method="loess") +
  geom_point(shape=18) +
  ylim(-10, 25) +
  labs(y = "",
       x = "",
       # y = "S. viride \nFeeding rate (bites/min)", 
       # x = expression(atop(Scarid~biomass, "(kg/100m"^2*")")),
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        plot.margin = margin(1, 1, 0, 1, unit = "pt")
        )

s2 <- ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), aes(x = consp_den, y = for_bites_mean)) +
  geom_smooth(method="loess") +
  geom_point(shape=18) +
  ylim(-10, 25) +
  labs(y = "", 
       x = "",
       # x = expression(atop(Conspec.~density, "(indv./100m"^2*")")),
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(1, 1, 0, 1, unit = "pt")
        )

s3 <- ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), aes(x = rugosity, y = for_bites_mean)) +
  geom_smooth(method="loess") +
  geom_point(shape=18) +
  ylim(-10, 25) +
  labs(y = "", 
       x = "",
       #x = expression("Rugosity"), 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(1, 1, 0, 1, unit = "pt")
        )

s4 <- ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), aes(x = ta_canopy, y = for_bites_mean)) +
  geom_smooth(method="loess") +
  geom_point(shape=18) +
  ylim(-10, 25) +
  labs(y = "", 
       x = "",
       # x = expression("EAM canopy \nheight (mm)"), 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(1, 1, 0, 1, unit = "pt")
        )

s5 <- ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), aes(x = ta_cover, y = for_bites_mean)) +
  geom_smooth(method="loess") +
  geom_point(shape=18) +
  ylim(-10, 25) +
  labs(y = "", 
       x = "",
       # x = expression("EAM percent \ncover"), 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(1, 1, 0, 1, unit = "pt")
        )

s7 <- ggplot(filter(sum_ssp, species_code == "stop" & phase == "i"), aes(x = ma_cover, y = for_bites_mean)) +
  geom_smooth(method="loess") +
  geom_point(shape=18) +
  ylim(-10, 25) +
  labs(y = "", 
       x = "",
       # x = expression("Macroalgal \npercent cover"), 
       title = "") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        plot.margin = margin(1, 1, 0, 1, unit = "pt")
        )

grid.arrange(left = "Feeding intensity (bites / foray)",
  arrangeGrob(q1, q3, q4, q5, q7, ncol=5, right = "S. vetula"),
  arrangeGrob(
    arrangeGrob(s1, bottom = "Scarid biomass"), 
    arrangeGrob(s3, bottom = "Rugosity"),
    arrangeGrob(s4, bottom = "Turf height"),
    arrangeGrob(s5, bottom = "Turf % cover"),
    arrangeGrob(s7, bottom = "Macro. % cover"),
    ncol=5, right = "S. viride"), 
  nrow=2) 
```

## PCA
```{r, echo=F, message=F, warning=F}
pca_pred <- prcomp(na.omit(sum_site %>% 
              select(rugosity, ma_cover, ma_canopy, ta_cover, ta_canopy, scar_den, 
                     scar_bm, carn30_bm, spear_daily)),
  center = T, scale = T)

autoplot(pca_pred, data = sum_site, colour = 'island',
         # frame = T, frame.type = "norm",
         loadings = T, loadings.colour = 'gray',
         loadings.label = T, loadings.label.size = 3, loadings.label.colour = 'black',
            loadings.label.repel = T) +
    scale_color_manual(name="island", values=c("lightblue", "darkseagreen", "lightsalmon")) + 
    theme(legend.direction = 'horizontal', legend.position = 'top') +
    theme_minimal()
ggsave(here("figs/pca_pred.png"))

sum_site <- sum_site %>% 
  bind_cols(as.data.frame(pca_pred$x) %>% 
           select(PC1,PC2)) %>%
  rename(pc1_all=PC1,
         pc2_all=PC2)

sum_id_pca <- sum_id %>% 
  left_join(select(sum_site,site,pc1_all,pc2_all), by = "site")
sum_ssp_pca <- sum_ssp %>%
  left_join(select(sum_site,site,pc1_all,pc2_all), by = "site")
```
```{r, eval = F, echo=F, message=F, warning=F}
ggplot(filter(sum_ssp_pca, species_code != "rbp" & phase == "i"), aes(x = pc1_all, y = fr_mean/60)) +
  facet_wrap(.~species) +
  geom_smooth(method="loess") +
  geom_point(shape=18) +
  labs(y = "Feeding rate (bites/min)", 
       x = "PC1")
  theme_bw()
```




## GAMM results

```{r, eval=F, echo=F, message=F, warning=F}
m_data <- sum_ssp_pca %>% filter(species_code!="rbp")
# Main effects only
gamm1 <- gamm(fr_mean ~ species_code + s(length_m) + phase + s(pc1_all) + s(pc2_all), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
summary(gamm1$gam)
plot(gamm1$gam,pages=1)
plot(gamm1$lme) # not too heteroskedastic?

# Length/species and PC1/species interactions
gamm2 <- gamm(fr_mean ~ species_code + phase + s(length_m, by = factor(species_code)) + s(pc1_all, by = factor(species_code)) + s(pc2_all), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
summary(gamm2$gam)
plot(gamm2$gam,pages=1)
plot(gamm2$lme)

# Length/species and PC1/species interactions, no PC2
gamm3 <- gamm(fr_mean ~ species_code + phase + s(length_m, by = factor(species_code)) + s(pc1_all, by = factor(species_code)), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
summary(gamm3$gam)
plot(gamm3$gam,pages=1)
plot(gamm3$lme) 

AIC(gamm1$lme, gamm2$lme, gamm3$lme)
AICc(gamm1$lme)
AICc(gamm2$lme)
AICc(gamm3$lme)
```

```{r, echo=F, message=F, warning=F}
m_data <- sum_ssp_pca %>% filter(species_code!="rbp")
# Length/species and PC1/species interactions
gamm2 <- gamm(fr_mean ~ species_code + phase + s(length_m, by = factor(species_code)) + s(pc1_all, by = factor(species_code)) + s(pc2_all), 
           family = gaussian(link=identity),
           random = list(island=~1),
           method = "REML",
           data = m_data
           )
summary(gamm2$gam)
plot(gamm2$gam,pages=1)
```


# Discussion

* GAMM results: explain meaning/significance of PC1. *S. vetula* seem more impacted by PC1 than *S. viride* 
* Potential mechanisms/hypotheses for these observed trends:
    * Social feeding hypothesis: more time must be spent vigilant in smaller groups, potential compounding effect with overfishing/loss of herbivory
    * Nutrient limitation hypothesis: scarids must graze more (and more intensely) in sites with lower algal canopy heights and percent cover
    * Predation risk hypothesis: presence of large piscivores does not explain feeding trends, but risk from spearfishers may play a bigger role (able to target larger fish, i.e. those sampled here). (Incorporate rugosity here?)
* Caveats + limitations: all these different factors are understandably correlated in observational work
    * room for experimental work to unpack/distinguish drivers
    * room (/need) for incorporation of behavior in coral reef grazing models, especially those used to set targets for minimum herbivore biomass levels, because these may look very different in different reef conditions.
* Conservation + mgmt implications: multiple human impacts on reefs may affect herbivore behavior in different ways, spec. fishing + habitat degradation - which may have reinforcing/compounding negative effects on reef health
